### 2012-11-25 paper.tex
replaced the todo addressed to my by merging preprocessed subsection and first decision tree section

\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute.

\paragraph{}The last preprocessing step was the replacement of missing values (data imputation). Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables. One data record did contain completeley useless data and was thus removed (see at in listing \ref{lst:crappy-entry}).

@@

\subsection{Decision Tree}
\todo[inline]{robert: do you want to keep this section? on the one hand it shows what happens if we run a desicion tree with attributes on which the the prediction attribute depends, but on the other hand we know we should not do that. please decide. i'm ok with both (philipp)}
%------------------------------------------------
On the first glimpse, it is an easy task for a decision based algorithm:

\begin{lstlisting}
J48 pruned tree
------------------
survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{lstlisting}

\paragraph{}That's correct, if the patient survived $ <= 12 $ months, he survived at least 12 month. What a statement!

\paragraph{}The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?). E.g. choosing a DecisionStump results in the expected single splitting: 
\begin{lstlisting}
Class distributions
survival <= 11.5
0   1   
1.0 0.0 
survival > 11.5
0   1   
0.0 1.0 
survival is missing
0   1   
0.042105263157894736    0.9578947368421052
\end{lstlisting}
\paragraph{}Since there is no missing value for the survival attribute, this branch can be ignored.

\paragraph{}But allowing the machine learning algorithms to see the origin attributes from which we derived the prediction attribut leads to absurdity, since the actual prediction task is to say whether a patient will survive one year after suffering a heart attack or not \emph{depending on the medical characteristics}.

\subsection{Decision Tree on preprocessed data}

\paragraph{}The J48 Decision Tree shows a more useful tree on the preprocessed data:

>>>>

\paragraph{}It should be clear that the variables ``survival'' and ``still-alive'' should not be used for classification since they constitute the final prediction attribute. Listing \ref{lst:silly-tree} shows what happens if a decision tree can take these attributes into account. It states that if the patient survived $ <= 12 $ months, he survived at least 12 month. Pretty accurate, right? 
The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute is redundant since $ x > 11 \rightarrow x > 12 $.

\begin{lstlisting}[caption={Decision tree on 100\% dependant attributes}, label=lst:silly-tree, float=htb]
J48 pruned tree
------------------
    survival <= 12
    |   survival <= 11: 0 (4.0)
    |   survival > 11: 1 (4.0)
    survival > 12: 1 (87.0)
\end{lstlisting}

\paragraph{}The last preprocessing step was the replacement of missing values (data imputation). Most of the data was complete but there were still some data records having missing values which we replaced by the mean values of the specific variables. One data record did contain completeley useless data and was thus removed (see in listing \ref{lst:crappy-entry}).

@@

\subsection{Decision Tree}

The J48 Decision Tree shows a more useful tree on the preprocessed data:
